Solving Collective Spin Hamiltonians with Decoherence

### 2018-08-20 soc_squeezing/cumulant_expansions.tex
solved last commutator!

\newcommand{\C}{\mathcal{C}}
\newcommand{\D}{\mathcal{D}}
\renewcommand{\O}{\mathcal{O}}

@@

  &= n \sum_{\ell=0}^{m-1} S_\mu^{m-\ell-1} S_\bmu^{n-1}
  \p{1 - n + 2\mu S_z} S_\mu^\ell + S_\bmu^n S_\mu^m,

@@

  \equiv n \sum_{\ell=0}^{m-1} S_\mu^{m-\ell-1} S_\bmu^{n-1}
  \p{1 - n + 2\mu S_z} S_\mu^\ell.

@@

We begin by pushing $S_\mu^\ell$ to the left of $S_z$, which takes
$2\mu S_z\to 2\mu S_z+2\ell$, and then push $S_\mu^{m-\ell-1}$ to the
right of $S_\bmu^{n-1}$, giving us

@@

  &= n \sum_{\ell=0}^{m-1}
  \p{S_\bmu^{n-1} S_\mu^{m-\ell-1} + C_{\mu,m-\ell-1,n-1}} S_\mu^\ell
  \p{2\ell + 1 - n + 2\mu S_z} \\

@@

  + n \sum_{\ell=0}^{m-2} C_{\mu,m-\ell-1,n-1}
  S_\mu^\ell \p{2\ell + 1 - n + 2\mu S_z},

@@

$C_{\mu,m-\ell-1,n-1}=\sp{S_\mu^{m-\ell-1},S_\bmu^{n-1}}_-=0$ if
$\ell=m-1$, and

@@

to order the spin operators, as $C_{\mu,m-\ell-1,n-1}$ contains $S_z$
operators which are to the left of $S_\mu^\ell$.  To sort all spin

@@

  C_{\mu mn}^{(\ell)} \equiv C_{\mu,m-\ell,n} S_\mu^\ell,

@@

  D_{\mu mn}^{(\ell)} \equiv D_{\mu,m-\ell,n} S_\mu^\ell,

@@

  D_{\mu mn}^{(\ell)}
  &\equiv D_{\mu,m-\ell,n}S_\mu^\ell \\
  &= \p{m-\ell}n S_\bmu^{n-1} S_\mu^{m-\ell-1}
  \p{m-\ell-n+2\mu S_z} S_\mu^\ell \\
  &= \p{m-\ell}n S_\bmu^{n-1} S_\mu^{m-1} \p{\ell+m-n+2\mu S_z},
  \label{eq:D_mn_ell}

@@

  C_{\mu mn}^{(\ell)}
  &= D_{\mu,m-\ell,n} S_\mu^\ell + n \sum_{k=0}^{m-\ell-2}
  C_{\mu,m-\ell-k-1,n-1} S_\mu^k \p{2k+1-n+2\mu S_z} S_\mu^\ell \\
  &= D_{\mu mn}^{(\ell)} + n \sum_{k=0}^{m-\ell-2}
  C_{\mu,m-\ell-k-1,n-1} S_\mu^{k+\ell} \p{2k+2\ell+1-n+2\mu S_z} \\
  &= D_{\mu mn}^{(\ell)} + n \sum_{k=0}^{m-\ell-2}
  C_{\mu,m-1,n-1}^{(\ell+k)} \p{2\sp{k+\ell}+1-n+2\mu S_z} \\
  &= D_{\mu mn}^{(\ell)} + n \sum_{k=\ell}^{m-2}
  C_{\mu,m-1,n-1}^{(k)} \p{2k+1-n+2\mu S_z}.
  \label{eq:C_mn_ell}

@@

in \eqref{eq:C_mn_ell} have been ordered.  We can now repeatedly
substitute $C_{\mu mn}^{(\ell)}$ into itself, each time decreasing $m$

@@

  C_{\mu mn}^{(\ell)}
  = D_{\mu mn}^{(\ell)}
  + \sum_{p=1}^{\min\set{m,n}-1} E_{\mu mn}^{(\ell,p)},

@@

where the first terms in the sum are

@@

  E_{\mu mn}^{(\ell,1)}
  = n \sum_{k=\ell}^{m-2} D_{\mu,m-1,n-1}^{(k)}

@@

@@

@@

  E_{\mu mn}^{(\ell,2)}
  = n \sum_{k_1=\ell}^{m-2} \p{n-1} \sum_{k_2=k_1}^{m-3}

@@

and more generally

@@

  E_{\mu mn}^{(k_0,p)}
  = n^{\underline{p}} \sum_{\p{\v k,k_0,m,p}}
  D_{\mu,m-p,n-p}^{(k_p)} \prod_{j=1}^p \p{2k_j+j-n+2\mu S_z}.
  \label{eq:E_mn_kp}

@@

with
$n^{\underline{p}}\equiv
n\p{n-1}\cdots\p{n-p+1}=\prod_{j=0}^{p-1}\p{n-j}$ and

@@

  \sum_{\p{\v k,k_0,m,p}}
  \equiv \sum_{k_1=k_0}^{m-2} \cdots \sum_{k_p=k_{p-1}}^{m-p-1}.

@@

Using \eqref{eq:D_mn_ell} to substitute $D_{\mu,m-p,n-p}^{(k_p)}$ into
\eqref{eq:E_mn_kp} gives us

@@

  E_{\mu mn}^{(k_0,p)}
  &= n^{\underline{p}} S_\bmu^{n-p-1} S_\mu^{m-p-1} \tag*{} \\
  &\qquad \times \sum_{\p{\v k,k_0,m,p}}
  \p{m-p-k_p} \p{n-p} \p{k_p+m-n+2\mu S_z} \tag*{} \\
  &\qquad \times \prod_{j=1}^p \p{2k_j+j-n+2\mu S_z}.
We now wish to pick off the coefficients of each operator in
$E_{\mu mn}^{\p{k_0,p}}$, so we express
  \prod_{j=1}^p \p{2k_j+p-n+2\mu S_z}
  = \sum_{q=0}^p \p{2\mu S_z}^q
  \sum_{c\in\C\p{q,p}} \prod_{\alpha\notin c} \p{2k_\alpha+\alpha-n},

@@

where implicitly $\alpha\in\set{1,2,\cdots,p}$, and $\C\p{q,p}$
denotes the set of $q$-combinations (i.e.~subsets with $q$ elements)
of $\set{1,2,\cdots,p}$; the number of elements in $\C\p{q,p}$ is
${p \choose q}$.  Note that for the expansion in
\eqref{eq:product_expansion} to hold, we require that the empty
product
$\prod_{\alpha\notin\set{1,\cdots,p}}\p{2k_\alpha+\alpha-n}=1$.  We
can now write

@@

  E_{\mu mn}^{\p{k_0,p}}
  &= n^{\underline{p}} S_\bmu^{n-p-1} S_\mu^{m-p-1}
  \sum_{q=0}^p \p{2\mu}^q S_z^q \tag*{} \\
  &\qquad \times \sum_{c\in\C\p{q,p}} \sum_{\p{\v k,k_0,m,p}}
  \p{m-p-k_p} \p{n-p} \p{k_p+m-n+2\mu S_z}
  \prod_{\alpha\notin c} \p{2k_\alpha+\alpha-n}.

@@

Solving \eqref{eq:C_mn_ell} may be more trouble than it's worth
right now, but to derive a system of equations for the two-axis
twisting Hamiltonian we need to know $\sp{S_\mu^2,S_\bmu^n}_-$, so we
expand
\begin{align}
  \sp{S_\mu^2,S_\bmu^n}_-
  = C_{\mu mn}
  = D_{\mu,2,n} + n D_{\mu,1,n-1} \p{1 - n + 2\mu S_z}
\end{align}

>>>>

\newcommand{\C}{\mathcal{C}}
\newcommand{\D}{\mathcal{D}}
\renewcommand{\O}{\mathcal{O}}
\renewcommand{\S}{\mathcal{S}}

\newcommand{\Z}{\mathbb{Z}}

@@

  &= n \sum_{k=0}^{m-1} S_\mu^{m-k-1} S_\bmu^{n-1}
  \p{1 - n + 2\mu S_z} S_\mu^k + S_\bmu^n S_\mu^m,

@@

  \equiv n \sum_{k=0}^{m-1} S_\mu^{m-k-1} S_\bmu^{n-1}
  \p{1 - n + 2\mu S_z} S_\mu^k.

@@

We begin by pushing $S_\mu^k$ to the left of $S_z$, which takes
$2\mu S_z\to 2\mu S_z+2k$, and then push $S_\mu^{m-k-1}$ to the right
of $S_\bmu^{n-1}$, giving us

@@

  &= n \sum_{k=0}^{m-1}
  \p{S_\bmu^{n-1} S_\mu^{m-k-1} + C_{\mu,m-k-1,n-1}} S_\mu^k
  \p{2k + 1 - n + 2\mu S_z} \\

@@

  + n \sum_{k=0}^{m-2} C_{\mu,m-k-1,n-1}
  S_\mu^k \p{2k + 1 - n + 2\mu S_z},

@@

$C_{\mu,m-k-1,n-1}=\sp{S_\mu^{m-k-1},S_\bmu^{n-1}}_-=0$ if $k=m-1$,

@@

to order the spin operators, as $C_{\mu,m-k-1,n-1}$ contains $S_z$
operators which are to the left of $S_\mu^k$.  To sort all spin

@@

  C_{\mu mn}^{(k)} \equiv C_{\mu,m-k,n} S_\mu^k,

@@

  D_{\mu mn}^{(k)} \equiv D_{\mu,m-k,n} S_\mu^k,

@@

  D_{\mu mn}^{(k)}
  &\equiv D_{\mu,m-k,n}S_\mu^k \\
  &= \p{m-k}n S_\bmu^{n-1} S_\mu^{m-k-1}
  \p{m-k-n+2\mu S_z} S_\mu^k \\
  &= \p{m-k}n S_\bmu^{n-1} S_\mu^{m-1} \p{k+m-n+2\mu S_z},
  \label{eq:D_mn_k}

@@

  C_{\mu mn}^{(k)}
  &= D_{\mu,m-k,n} S_\mu^k + n \sum_{j=0}^{m-k-2}
  C_{\mu,m-k-j-1,n-1} S_\mu^j \p{2j+1-n+2\mu S_z} S_\mu^k \\
  &= D_{\mu mn}^{(k)} + n \sum_{j=0}^{m-k-2}
  C_{\mu,m-k-j-1,n-1} S_\mu^{j+k} \p{2j+2k+1-n+2\mu S_z} \\
  &= D_{\mu mn}^{(k)} + n \sum_{j=0}^{m-k-2}
  C_{\mu,m-1,n-1}^{(k+j)} \p{2\sp{j+k}+1-n+2\mu S_z} \\
  &= D_{\mu mn}^{(k)} + n \sum_{j=k}^{m-2}
  C_{\mu,m-1,n-1}^{(j)} \p{2j+1-n+2\mu S_z}.
  \label{eq:C_mn_k}

@@

in \eqref{eq:C_mn_k} have been ordered.  We can now repeatedly
substitute $C_{\mu mn}^{(k)}$ into itself, each time decreasing $m$

@@

  = C_{\mu mn}^{(0)}
  = D_{\mu mn}
  + \sum_{p=1}^{\min\set{m,n}-1} E_{\mu mn}^{(p)},

@@

where the first terms in the sum over $p$ is

@@

  E_{\mu mn}^{(1)}
  = n \sum_{k=0}^{m-2} D_{\mu,m-1,n-1}^{(k)}

@@

  \label{eq:E_mn_1_first}

@@

the second

@@

  E_{\mu mn}^{(2)}
  = n \sum_{k_1=0}^{m-2} \p{n-1} \sum_{k_2=k_1}^{m-3}

@@

and more generally for $p>1$,

@@

  E_{\mu mn}^{(p)}
  = n^{\underline{p}}
  \sum_{k_1=0}^{m-2}\cdots\sum_{k_p=k_{p-1}}^{m-p-1}
  D_{\mu,m-p,n-p}^{(k_p)} \prod_{j=1}^p \p{2k_j+j-n+2\mu S_z},

@@

with $n^{\underline{p}}\equiv n\p{n-1}\cdots\p{n-p+1}$.  Observing
that we always have $k_1\le k_2\le\cdots\le k_p$, we can rearrange the
sums to get that for $p>1$

@@

  E_{\mu mn}^{(p)}
  = n^{\underline{p}}
  \sum_{\ell=0}^{m-p-1} D_{\mu,m-p,n-p}^{(\ell)} \p{2\ell+p-n+2\mu S_z}
  \sum_{\substack{\v k\in\Z_{\ell+1}^{p-1}\\k_1\le k_2\le\cdots\le k_{p-1}}}
  \prod_{j=1}^{p-1} \p{2k_j+j-n+2\mu S_z},
  \label{eq:E_mn_p_semifinal}

@@

where $\Z_q$ denotes the set of $q$ smallest non-negative integers,
e.g.~$\Z_{\ell+1}=\set{0,1,\cdots,\ell}$.  We now expand

@@

  \prod_{j=1}^{p-1} \p{2k_j+j-n+2\mu S_z}
  = \sum_{q=0}^{p-1} \p{2\mu S_z}^q
  \sum_{c\in\C_{p-1-q}\p{\Z_{p-1}}} \prod_{j\in c} \p{2k_{j+1}+j+1-n},

@@

where $\C\p{k,\S}$ denotes the set of $k$-combinations (i.e.~subsets
with $k$ elements) of $\S$; the number of elements in
$\C_{p-1-q}\p{\Z_{p-1}}$ is
${ p-1 \choose p-1-q } = { p-1 \choose q }$.  Substituting
\eqref{eq:D_mn_k} and \eqref{eq:product_expansion} into
\eqref{eq:E_mn_p_semifinal}, we get that for $p>1$,

@@

  E_{\mu mn}^{(p)}
  = n^{\underline{p}} S_\bmu^{n-p-1} S_\mu^{m-p-1}
  \sum_{s=0}^2 f_{mnp\ell}^{(s)} \p{2\mu S_z}^s
  \sum_{q=0}^{p-1} g_{np\ell q} \p{2\mu S_z}^q,
  \label{eq:E_mn_p}
where
  f_{mnp\ell}^{(0)}
  &= \p{m-p-\ell} \p{n-p} \p{\ell+m-n} \p{2\ell-n+p}, \\
  f_{mnp\ell}^{(1)}
  &= \p{m-p-\ell} \p{n-p} \p{3\ell+m-2n+p}, \\
  f_{mnp\ell}^{(2)}
  &= \p{m-p-\ell} \p{n-p}, \\
  g_{np\ell q}
  &=
  \sum_{c\in\C_{p-1-q}\p{\Z_{p-1}}}
  \sum_{\substack{\v k\in\Z_{\ell+1}^{p-1}\\k_1\le k_2\le\cdots\le k_{p-1}}}
  \prod_{j\in c} \p{2k_{j+1}+j+1-n},
and for $p=1$ we can resolve \eqref{eq:E_mn_1_first} to get
  E_{\mu mn}^{(1)}
  = \f16 mn \p{m-1} \p{3n-m-1} S_\bmu^{n-1} S_\mu ^{m-1} \p{m+2\mu S_z}.
  \label{eq:E_mn_1}
With reference to \eqref{eq:E_mn_p}-\eqref{eq:E_mn_1}, we finally have
  = mn S_\bmu^{n-1} S_\mu^{m-1}
  \sum_{q\in\set{0,1}} \p{m-n}^{1-q}\p{2\mu S_z}^q
  + \sum_{p=1}^{\min\set{m,n}-1} E_{\mu mn}^{(p)}.

@@

To derive a system of equations for the two-axis twisting Hamiltonian
we need to know $\sp{S_\mu^2,S_\bmu^n}_-$, so we expand
\begin{align}
  \sp{S_\mu^2,S_\bmu^n}_-
  = C_{\mu mn}
  = D_{\mu,2,n} + n D_{\mu,1,n-1} \p{1 - n + 2\mu S_z}
\end{align}
